---
title: "EMAIL CLASSIFICATION"
output: pdf_document
date: "2023-04-17"
---

## DATA LOADING
```{r}
data=read.csv("spam_ham_dataset.csv")
names(data)=c("col","label","text","label_num")

```

## DATA PREPROCESSING

```{r}
#Check the overview of the data
head(data)
```

```{r}
#Check for any missing values
sum(is.na(data))
```

There are no missing values so there is no need of deleting any rows.

```{r}
library(tidyverse)
```


```{r}
#Plot a boxplot to check whether there are any outliers in the dataset.
library(ggplot2)
ggplot(data, aes(y = data$col)) +
  geom_boxplot()
```

There are no outliers from the dataset. Now we will check the datatypes of the columns.
```{r}
#Check the datatypes of each column.
str(data)
```

Yes, all the columns are having the correct datatypes. Because if we see the label it should be categorical but a separate column has been created to represent it as a category. That is "label_num".


```{r}
data$text<-gsub('[^[:alnum:] ]', ' ',data$text )
data$text<-gsub('Subject','',data$text)
head(data)
```
The text attribute is modified to remove punctuation marks and other special characters within the text. Also the subject tag is removed
as it remains the same. This helps identify key words in classification.



```{r}
library(tm)
library(e1071)

dat_corpus <- Corpus(VectorSource(data$text))

# Clean the corpus
dat_corpus <- tm_map(dat_corpus, tolower)
dat_corpus <- tm_map(dat_corpus, removeNumbers)
dat_corpus <- tm_map(dat_corpus, removePunctuation)
dat_corpus <- tm_map(dat_corpus, stripWhitespace)
dat_corpus <- tm_map(dat_corpus, removeWords, stopwords("english"))
dat_dtm <- DocumentTermMatrix(dat_corpus)
dat_dtm
```
Removing terms which don't occur frequently
```{r}
sdtm<-removeSparseTerms(dat_dtm,0.95)
sdtm
```
Converting the word sparse matrix to a dataframe 
```{r}
word_sparse=data.frame(as.matrix(sdtm),email_class=data$label)
head(word_sparse)
```

## DATA SPLITTING

```{r}
library(caret)
# Set seed
set.seed(123)



# Splitting the dataset into training and testing sets

train_index <- createDataPartition(word_sparse$email_class, p = 0.7, list = FALSE)
train_data <- word_sparse[train_index, ]
test_data <- word_sparse[-train_index, ]

```

```{r}
train_data$email_class <- factor(train_data$email_class, levels = c("spam", "ham"))
test_data$email_class <- factor(test_data$email_class, levels = c("spam", "ham"))
```

## BUILDING NAIVE BAYES CLASSIFICATION MODEL
```{r}
# Building the Naive Bayes Classifier Model
library(e1071)
nb_model <- naiveBayes(email_class ~ ., data = train_data)
```

```{r}
# Use the classifier to make predictions on the test data
predicted_labels = predict(nb_model, newdata = test_data[,-ncol(test_data)])
```

Our model has been built. Now we will check the accuracy of the model.
```{r}
levels(predicted_labels)
```
```{r}
levels(test_data$email_class)

```

```{r}
library(caret)
confusionMatrix(predicted_labels, test_data$email_class)
```
The accuracy of the model is 70%.

# Logistic Regression Model


```{r}
X_train <- data.matrix(train_data[, -ncol(train_data)])
Y_train <- (train_data$email_class)
X_test <- data.matrix(test_data[, -ncol(test_data)])
Y_test <- (test_data$email_class)
```

Training the logistic regression model
```{r}
library(glmnet)
lr<-cv.glmnet(X_train,Y_train, family = "binomial")
```

Predicting test data
```{r}
prediction_lr<-predict(lr,newx=X_test,type="class")
```

Accuracy details
```{r}
library(caret)
confusionMatrix(as.factor(prediction_lr),as.factor(Y_test))
```